NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-06-07 Manuscript/manuscriptPS.tex
Update text

The headgroup order parameters of DOPS and POPS bilayers 
from different simulation models are compared with the experimental
data in Fig. \ref{HGorderParametersPS}. In line with the previous
study for PC lipids \cite{botan15}, the glycerol backbone order 
parameters in CHARMM36 roughly agree with the experimental data,
while significant discrepancies for glycerol backbone carbons
are observed in other models.
\todo{Discussion is to be finished.} \\
\todo{Also the discussion about POPS/OPPS issue with MacRog model should be added.}

Based on the subjective ranking shown in Fig. \ref{comparisonTablePS},
the CHARMM36 is the best performing model for both lipids.
However, none of the tested models give a satisfactory 
agreement with experiments for the headgroup order paramaters of 
PS lipids. The total deviation from the experiments ($\Sigma$ in Fig. \ref{comparisonTablePS})
for the best performing CHARMM36 model is larger for PS lipids (8)
than for PC (3) \cite{botan15}. Therefore the 
interpretation of structural details of PS headgroup or
differences between PC and PS lipids is very challenging 
from the current MD simulation models.
Figures \ref{dihedralsGLY} and \ref{dihedralsHG} show dihedral
angle distributions calculated from different models. 
The glycerol backbone structures from CHARMM36 and
Slipids simulations visualized in Fig. \ref{glycerol_buslaev} reveal the
missing structures in Slipids model, which probably lead to the incorrect
order parameters. For the headgroup $\alpha$ and $\beta$ carbon order parameters of PS lipids,
the tested models perform less well than for PC headgroup in previous
study \cite{botan15}.
\todo{Discussion is to be finished.
  One possible conclusion could be the following:
  The main differences between the models in the headgroup region are observed
  for dihedrals C12-C11-O12-P and C11-C12-C13-O1A.
  CHARMM36, CHARMM36UA and Slipids give very similar results to the
  dihedral C11-C12-C13-O1A, which is close to the $\beta$-carbon.
  The order parameters of $\beta$-carbons for these three models
  are in best argeement with the experiments in figure \ref{HGorderParametersPS}.
  On the other hand, Gromos-CKP models give better order parameters
  for $\alpha$-carbon than Slipids, CHARMM36 or CHARMM36UA. 
  In conclusion, the suggestion would be that the single peak for
  observed at 120 degrees in CHARMMs and Slipids would
  be more realistic for C11-C12-C13-O1A dihedral, while the single peak at 180 degrees observed
  in CKP models and in Berger would be most realistic for C12-C11-O12-P dihedral.
}

@@

>>>>

The headgroup order parameters of DOPS and POPS bilayers 
from different simulation models are compared with the experimental
data in Fig. \ref{HGorderParametersPS}. Subjective ranking
of the quality of the models is shown in Fig. \ref{comparisonTablePS}.
The tested models perform generally less well than in the previous
study for PC headgroup~\cite{botan15} and none of the models 
reproduce the experimental order parameters within
experimental error bars.
Therefore, the models cannot be straightforwardly used to interpret the
structure of PS headgroup. However, the differences between PC and PS headgroups
are partially reproduced by some of the models.

The two best performing models for the $\alpha$ and $\beta$-carbons of
PS, Slipids and CHARMM36, reproduce the large forking in the $\alpha$-carbon
and the Slipids model gives also a good agreement with experiments for
the $\beta$-carbon order parameter in both PC and PS headgroups
(Fig. \ref{HGorderParametersPS} and Ref.~\citenum{botan15}).
Interestingly, the dihedral angle distributions of CHARMM36
and Slipids in Fig.~\ref{dihedralsHG} share significant similarities
in the headgroup region. However, the experimental order parameters in the
glycerol backbone region are not well reproduced by the Slipids model,
which was also the case in for PC lipids \cite{botan15}.
This difference probably arises from the differences in
the dihedral angle distributions of C1-C2-C3-O31 and C2-C3-O31-C31
in Figure \ref{dihedralsGLY}, which are also illustrated in
Figure \ref{glycerol_buslaev}.

%\todo{Discussion is to be finished.
%  One possible conclusion could be the following:
%  The main differences between the models in the headgroup region are observed
%  for dihedrals C12-C11-O12-P and C11-C12-C13-O1A.
%  CHARMM36, CHARMM36UA and Slipids give very similar results to the
%  dihedral C11-C12-C13-O1A, which is close to the $\beta$-carbon.
%  The order parameters of $\beta$-carbons for these three models
%  are in best argeement with the experiments in figure \ref{HGorderParametersPS}.
%  On the other hand, Gromos-CKP models give better order parameters
%  for $\alpha$-carbon than Slipids, CHARMM36 or CHARMM36UA. 
%  In conclusion, the suggestion would be that the single peak for
%  observed at 120 degrees in CHARMMs and Slipids would
%  be more realistic for C11-C12-C13-O1A dihedral, while the single peak at 180 degrees observed
%  in CKP models and in Berger would be most realistic for C12-C11-O12-P dihedral.
%}

\todo{Also the discussion about POPS/OPPS issue with MacRog model should be added.}

@@

%
%The total deviation from the experiments ($\Sigma$ in Fig. \ref{comparisonTablePS})
%for the best performing CHARMM36 model is larger for PS lipids (8)
%than for PC (3) \cite{botan15}.
%Therefore the interpretation of structural details of PS headgroup or
%differences between PC and PS lipids is very challenging 
%from the current MD simulation models.
